Numerical simulation and entropy dissipative cure of the carbuncle instability for the shallow water circular hydraulic jump

### 2021-08-14 numerics_paper/version2/numerical_chj.tex

We model only the flow on the upstream side of the cylinder, since
our interest is in the resolution of the bow shock.  We take the domain
$[0,40]\times[0,100]$ with a cylinder of radius 20 centered at $(40,50)$.
Reflecting boundary conditions are imposed at the surface of the cylinder,
along with outflow conditions along the rest of the right edge of the domain.
The depth and velocity are set initially and (for all time) at the other
boundaries to $h_0=1$, $u_0=5$, for a Froude number of 5.
We use a $160 \times 400$ uniform Cartesian grid.
%
In Figure \ref{fig:steady_outflow}, we show results at $t=80$, after the flow has reached
a steady state.  With Roe's solver, negative values (of the water height) are generated at an early
time, leading to failure of the solver.  Therefore, we show results only for
Rusanov's, HLLEMCC by \cite{kemm2014note}, and the blended solvers.
{\color{red}
  These three methods give carbuncle free results.
  It is worth noting that the solution with Rusanov's solver is the most dissipated solution. 
  Let 
  \begin{align}\label{res}
    \xi(t^{n+1}):=\frac{||h^{n+1}-h^n||_{\ell^2}}{||h^0||_{\ell^2}}
  \end{align}
  denote the residual based on the water height. In Figure \ref{fig:residuals_cylinder_2nd}, we plot $\xi(t)$ considering
  the three Riemann solvers. With HLLEMCC or the blended solver, the residual stagnates. 
  In Figure \ref{fig:residuals_cylinder_1st} we repeat the same experiment but using the first-order method \eqref{first-order_via_fluct} instead.
  In this case, the residual stagnates only when we consider the HLLEMCC solver.
  Clearly, the ability of the blended solver to converge to the steady state solution is affected only by the second-order corrections.
  In Figure \ref{fig:cylinder_blended}, we zoom close to the front of the shock and show the grid we use.
  Note that the grid is cartesian and that the shock is resolved within few cells. Therefore, the ability of the method to deliver
  carbuncle free solutions does not depend on the shock being aligned with a grid interface. 
}

@@

{\color{OliveGreen}
  We inspect further the behavior of the blended solver across the shock. 
  We are interested in the possible presence of an unphysical spike in the momentum.
  This numerical artifact might lead to a slow displacement of the position of the shock,
  which is known as slowly-moving shock anomaly; see for instance \cite{navas2019improved, zaide2012numerical}.
  In the left panels of Figure \ref{fig:evol_of_hu}, we show the time evolution of $hu(x,y=50)$ using the Rusanov and the blended solvers.
  Clearly, with both solvers we obtain the spike in $hu$. This behavior is expected due to the non-linear Hugoniot locus
  and the presence of intermediate states through the numerical shock \cite[Section 4.1]{navas2019improved}.

  In \cite{navas2019improved}, it is remarked that spurious oscillations in the transverse direction are introduced with Roe's solver.
  The authors observed these oscillations by plotting the time evolution of a slice along a fixed value of $x$.
  The authors conjectured that increasing dissipation of the shear waves reduces these oscillations.
  Rusanov's solver adds more dissipation (than the solvers they use) to the shear waves; therefore,
  it eliminates the spurious oscillations,
  as seen in the right-top panel of Figure \ref{fig:evol_of_hu}. Small oscillations are present with the blended solver.
  %
  Importantly, with the Rusanov and the blended solvers, the position of the shock appears to remain constant
  (once the steady state is achieved). This is suggested by the parabolic profile of the space-time evolution of $hu(x=18,y)$.
  To further confirm this, we run the simulation up to $t=160$ and plot in Figure \ref{fig:pos_of_shock} the location of the shock along $y=50$.
  Note that after $t\approx 55$, the location of the shock remains constant.

>>>>

We model only the flow on the upstream side of the cylinder, since
our interest is in the resolution of the bow shock.  We take the domain
$[0,40]\times[0,100]$ with a cylinder of radius 20 centered at $(40,50)$.
Reflecting boundary conditions are imposed at the surface of the cylinder,
along with outflow conditions along the rest of the right edge of the domain.
The depth and velocity are set initially and (for all time) at the other
boundaries to $h_0=1$, $u_0=5$, for a Froude number of 5.
We use a $160 \times 400$ uniform Cartesian grid.
%
In Figure \ref{fig:steady_outflow}, we show results at $t=80$, after the flow has reached
a steady state.  With Roe's solver, negative values (of the water height) are generated at an early
time, leading to failure of the solver.  Therefore, we show results only for
Rusanov's, HLLEMCC by \cite{kemm2014note}, and the blended solvers.
{\color{red}
  These three methods give carbuncle free results.
  It is worth noting that the solution with Rusanov's solver is the most dissipated solution. 
  Let 
  \begin{align}\label{res}
    \xi(t^{n+1}):=\frac{||h^{n+1}-h^n||_{\ell^2}}{||h^0||_{\ell^2}}
  \end{align}
  denote the residual based on the water height. In Figure \ref{fig:residuals_cylinder_2nd}, we plot $\xi(t)$ for the
  solution obtained with each Riemann solver. With HLLEMCC or the blended solver, the residual stagnates. 
  In Figure \ref{fig:residuals_cylinder_1st} we repeat the same experiment but using the first-order method \eqref{first-order_via_fluct} instead.
  We obtain similar results with Rusanov and HLLEMCC, but now the solution with the blended solver continues to converge
  at a rate similar to that of Rusanov.  This suggests that the lack of convergence for the 2nd-order method with the
  blended solver is due to the correction terms and not to the Riemann solver itself.
  In Figure \ref{fig:cylinder_blended}, we show a closeup of the front of the shock and show the grid we use.
  Note that the grid is cartesian and that the shock is resolved across a few
  cells. Therefore, the ability of the method to deliver
  carbuncle free solutions does not depend on the shock being aligned with a grid interface. 
}

@@

{\color{OliveGreen}
  In Figure \ref{fig:evol_of_hu} we show the time evolution of the momentum in the region around the shock.
  We see that the momentum is not monotone across the shock; this is a common feature of numerical shock
  solutions and in some cases is associated with a slow displacement of the position of the shock,
  which is known as slowly-moving shock anomaly; see for instance \cite{navas2019improved, zaide2012numerical}.
  This anomaly is, however, not present in the blended solver solution;
  the position of the shock appears to remain constant (once the steady state
  is achieved).
  To further confirm this, we run the simulation up to $t=160$ and plot in
  Figure \ref{fig:pos_of_shock} the location of the shock along $y=50$.
  Note that after $t\approx 55$, the location of the shock remains constant.

  Another numerical artifact that has been observed in this context is that of
  spurious oscillations in the transverse direction, behind the shock; these have been
  observed for instance when using Roe's solver \cite{navas2019improved}.
  Rusanov's solver adds more dissipation (than the solvers they use) to the shear waves; therefore,
  it eliminates the spurious oscillations,
  as seen in the right-top panel of Figure \ref{fig:evol_of_hu}. Small oscillations are present with the blended solver.
  %
